\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-12-18 popl-2019-journal-v/paper.tex
Pass over section 4.3

\end{jversion}
\subsection{Reasonable Components}
\label{sec:reasonable-components}
% Reasonability
The static guarantees given by $\wdjud[\ta]{\comp}$ make sure that components initially don't undermine the security measures needed for \stktokens{}, but it does not prevent a trusted component from doing something silly during execution that undermines \stktokens{}.
In order for \stktokens{} to provide guarantees for a component, we expect it not to shoot itself in the foot and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
More precisely, we expect four things of a reasonable component:
\begin{enumerate}[label=(\arabic*)]
\item It checks the stack base address before performing a call.
As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it often would be redundant.
\item It uses the return seals only for calls and the closure seals in an appropriate way which means that they should only be used to seal executable capabilities for code that behaves reasonably or non-executable things that do not undermine the security mechanisms \stktokens{} relies on.
\item It does not leak return and closure seals (or means to retrieve them).
This means that sets of seals with return or closure seals cannot be left in registers when transferring control to another module.
There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
\item It never stores return and closure seals (or means to get them) to memory.
By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them to begin with.
\end{enumerate}
\begin{jversion}
  We capture these properties in 4 definitions.
  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly or indirectly.
  Definition~\ref{def:reasonable-pc} defines the reasonable PCs which means that if it is plugged into a configuration with a register file filled with reasonable words, then the configuration behaves reasonably.
  Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
  Finally, definition~\ref{def:reasonable-component} lifts the notion of reasonability to components.

  \subsubsection{Reasonable words}
  To provide any guarantees, \stktokens{} rely on the program not to leak return seals in any way.
  But what does it mean to ``leak'' a return seal?
  It means that sets of seals that contain return seals as well as any means to obtain such sets cannot be leaked.
  The following definition makes this more precise.
\begin{definition}[Reasonable word]
  \label{def:reasonable-word}
  Take a set of trusted addresses $\ta$ and sets of return and closure seals $\gsigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
free stack $\ms_\stk$ if $n=0$
or the following implications hold.
  \begin{itemize}
  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then
    $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \mathrel{\#} \ta$
  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
    \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$
    and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all
    $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
  \item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$
    and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for
    all $\aaddr \in [\baddr,\eaddr]$
  \end{itemize}
\end{definition}
According to the definition, sets of seals that contain return or closure seals are not reasonable.
\stktokens{} rely on return seals to be unique in order to work, but it does not rely on closure seals.
However, it defeats the purpose of the closure seals if they are leaked to malicious code as it allows the malicious code to fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectually allows malicious code to unseal the data capability.
Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed with.

Whether a capability is reasonable depends on what it gives access to.
This is why the word reasonability is defined relative to the memory and stack.
For instance, if a read capability gives access to a piece of memory that contains a set-of-seals with a return seal, then it would not be reasonable.
This is why stack pointers and memory capabilities with read permission must only give access to memory with reasonable words.
The definition is cyclic, but the step-index ensures that it is well-founded.

@@

Therefore, it only seems natural to also define when a pc is reasonable.
Definition~\ref{def:reasonable-pc} roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file are reasonable and the result is a reasonable configuration.

@@

Item~\ref{item:dont-store-private} and Item~\ref{item:dont-leak-private} make sure that return seals are not leaked.
Item~\ref{item:dont-store-private} says that only reasonable words can be stored in memory.
This means that sets of return seals or execute capabilities cannot be stored in memory by a reasonable component.
Intuitively, a well-formed and reasonable component has its seals available in the code memory, so it can always retrieve them from the code memory.
In other words, it is not necessary to store them in the data memory.
Further, by making sure that seals are not stored in memory, we can allow capabilities for data memory to be handed out if there is a need for that (for instance to have a shared buffer).
Item~\ref{item:dont-leak-private} makes sure that seals are not leaked when transferring control to another component (i.e. on security boundary crossings).
With the component setup, there are two ways to transfer control: xjmp and call.
In both cases, we require that all of the argument registers contain reasonable words.
An execution configuration may need to do other operations than calling other code, and seals should no be leaked at any point during execution.
For this reason, Item~\ref{item:dont-leak-private} also says that if the next step is not a call or a jump, then the next execution configuration should also be reasonable.

@@

In our result, we assume that adversarial components are well-formed, but not necessarily reasonable. The well-formedness assumption ensures
that the trusted component can rely on basic security guarantees provided by the capability machine. For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
However, the adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
Can adversarial code call into trusted components?
The answer to that question is yes but not with LSE and WBCF guarantees.
Formally, adversarial code can contain the instructions that constitute a call.
However, for untrusted code, \srccm{} will not execute those instructions as a ``native call'' but execute the individual instructions separately.
The callee then executes in the same stack frame as the caller, so WBCF and LSE do not follow (for that call).

>>>>

\end{jversion}
\subsection{Reasonable Components}
\label{sec:reasonable-components}
% Reasonability
The static guarantees given by $\wdjud[\ta]{\comp}$ make sure that components initially don't undermine the security measures needed for \stktokens{}.
However, in order for \stktokens{} to provide guarantees for a component, we additionally expect it not to shoot itself in the foot at run-time and perform certain necessary checks not captured by the call code (Figure~\ref{fig:call-code}).
More precisely, we expect four things of a reasonable component:
\begin{enumerate}[label=(\arabic*)]
\item It checks the stack base address before performing a call.
As explained in Section~\ref{sec:stktokens-explained}, we do not include this check in the call code as it would often be redundant.
\item It uses the return seals only for calls and the closure seals in an appropriate way.
  Specifically, closure seals should only be used to seal executable capabilities for code that behaves reasonably, or for certain forms of non-executable capabilities.
\item It does not leak return and closure seals (or means to retrieve them).
This means that set-of-seals capabilities with return or closure seals cannot be left in registers when transferring control to another module.
There are also indirect ways to leak seals such as leaking a capability for code memory or leaking a capability for code memory sealed with an unknown seal.
\item It never stores return and closure seals (or means to get them) to memory.
By disallowing this, we make sure that data memory always can be safely shared as it does not contain seals or means to get them.
\end{enumerate}
\begin{jversion}
  We capture these properties in 4 definitions.
  Definition~\ref{def:reasonable-word} defines the reasonable words which means that they cannot be used to leak seals directly or indirectly.
  Definition~\ref{def:reasonable-pc} defines the reasonable PCs which means that if it is plugged into a configuration with a register file filled with reasonable words, then the configuration behaves reasonably.
  Definition~\ref{def:reasonable-conf} defines the reasonable configurations which captures the four informal behavioural properties.
  Finally, definition~\ref{def:reasonable-component} lifts the notion of reasonability to components.

  \subsubsection{Reasonable words}
  To provide any guarantees, \stktokens{} relies on the program not to leak return seals in any way.
  But what does it mean to ``leak'' a return seal?
  It means that set-of-seals capabilities that contain return seals as well as any means to obtain such sets cannot be leaked.
  The following definition makes this more precise.
\begin{definition}[Reasonable word]
  \label{def:reasonable-word}
  Take a set of trusted addresses $\ta$ and sets of return and closure seals $\gsigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ and
free stack $\ms_\stk$ if $n=0$
or the following implications hold.
  \begin{itemize}
  \item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then
    $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$, then $[\baddr,\eaddr] \mathrel{\#} \ta$
  \item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup
    \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
  \item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$
    and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all
    $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
  \item If $w = \stkptr{\perm,\baddr,\eaddr,\_}$ and $\perm \in \readAllowed{}$
    and $n > 0$, then $\ms_\stk(\aaddr)$ is reasonable up to $n - 1$ steps for
    all $\aaddr \in [\baddr,\eaddr]$
  \end{itemize}
\end{definition}
The definition rules out set-of-seals capabilities that contain return or closure seals.
\stktokens{} rely on return seals to be unique in order to work, but it does not rely on closure seals.
However, leaking closure seals to malicious code still defeats their purpose as the malicious code can then fabricate data capabilities for the code capabilities or code capabilities for the data capabilities which effectively allows malicious code to unseal the data capability.
Further, it makes reasoning about leakage of return seals difficult because the closures have to be well-behaved no matter what data they are executed with.
Capabilities to trusted code memory in $\ta$ are also ruled out, because the code memory may contain set-of-seals capabilities.
Capabilities that are sealed with capabilities not owned by the component must still be reasonable themselves (since such an untrusted seal provides no protection at all).

Finally, we define that stack or memory capabilities with read permission are reasonable if the memory they give access to contains only reasonable words.
To accomodate this, reasonability of words is defined relative to the memory and stack.
The definition is cyclic, but the step-index ensures that it is well-founded.

@@

Therefore, we define when an executable capability is reasonable as a pc.
The following definition roughly says that a pc is reasonable when you can plug it into a configuration where all the words in the register file are reasonable and the result is a reasonable configuration.

@@

Item~\ref{item:dont-store-private} and Item~\ref{item:dont-leak-private} make sure that return seals are not leaked.
Item~\ref{item:dont-store-private} says that only reasonable words can be stored in memory.
This means that sets of return seals or execute capabilities cannot be stored in memory by a reasonable component.
Intuitively, a well-formed and reasonable component has its seals available in the code memory, so it can always retrieve them from the code memory.
In other words, it is not necessary to store them in the data memory.
Further, by making sure that seals are not stored in memory, we can allow capabilities for data memory to be handed out if there is a need for that (for instance to have a shared buffer).
Item~\ref{item:dont-leak-private} makes sure that seals are not leaked when transferring control to another component (i.e. on security boundary crossings).
With the component setup, there are two ways to transfer control: xjmp and call.
In both cases, we require that all of the argument registers contain reasonable words.
An execution configuration may need to do other operations than calling other code, and seals should not be leaked at any point during execution.
For this reason, Item~\ref{item:dont-leak-private} also says that if the next step is not a call or a jump, then the next execution configuration should also be reasonable.

@@

In our result, we assume that adversarial components are well-formed.
This assumption ensures that the trusted component can rely on basic security guarantees provided by the capability machine.
For instance, if we did not require linearity to be respected initially, then adversarial code could start with an alias for the stack capability.
However, the adversary is not assumed to be reasonable as we do not expect them to obey the calling convention in any way.
Can adversarial code call into trusted components?
The answer to that question is yes but not with LSE and WBCF guarantees.
Formally, adversarial code can contain the instructions that constitute a call.
However, for untrusted code, \srccm{} will not execute those instructions as a ``native call'' but execute the individual instructions separately.
The callee then executes in the same stack frame as the caller, so WBCF and LSE do not follow (for that call).
